\bf Interactive Visualization of Hierarchically Structured Data

### 2017-09-22 treelapse.tex
some minor changes before resending

{\centering \includegraphics[width=300px]{figure/preterm_preterm}}

\caption{Samples with high levels of phyla other than Firmicutes appear to be
  related to preterm births. Here, we again display the preterm birth data with
  a DOI sankey, but instead of grouping samples according to statistically
  generated CSTs, we directly assign samples to preterm vs. not preterm
  according to whether the mother eventually had a preterm birth. These new
  states are visible in the updated legend. Through interactivity, it becomes
  possible to develop meaningful visual summaries even before calculating formal
  statistical ones.}
\label{fig:pretermpreterm}
\end{figure}

@@

\caption{An application to the Global Patterns demonstrates how linking in
  treelapse can be applied to combine hierarchical clustering and parallel
  coordinates views. Each panel represents a different subtree cluster within
  this data set, as indicated by the different locations for the tree brushes.
  The paths in the lower halves of each display represent the average value
  across different bacteria rather than timepoints, as in all previous figures.
  Though the samples originally do not include any hierarchical structure,
  hierarchical clustering provides such a structure which can then be
  interpreted using treeboxes. }\label{fig:gptimebox}
\end{figure}

In Figure \ref{fig:gptimebox}, we compare two subclusters from the
hierarchical clustering tree, after zooming to a few of the bacteria
that distinguish between the clusters. Upon revisiting the original
data, it becomes clear that the samples highlighted on the left come
from freshwater samples, while those on the right come from soil and
skin, and looking up taxonomic groups associated with the distinguishing
bacteria confirms this. For example, many of the species with high
abundances in the left figure come from order Oceanospirillales.

>>>>

\begin{figure}
\centering
\includegraphics[width=300px]{figure/preterm_preterm}
\caption{Samples with high levels of phyla other than Firmicutes appear to be
  related to preterm births. Here, we again display the preterm birth data with
  a DOI sankey, but instead of grouping samples according to statistically
  generated CSTs, we directly assign samples to preterm vs. not preterm
  according to whether the mother eventually had a preterm birth. These new
  states are visible in the updated legend. Through interactivity, it becomes
  possible to develop meaningful visual summaries even before calculating formal
  statistical ones.}
\label{fig:pretermpreterm}
\end{figure}

@@

\caption{An application to the Global Patterns demonstrates how linking in
  treelapse can be applied to combine hierarchical clustering and parallel
  coordinates views. Each panel represents a different subtree cluster within
  this data set, as indicated by the different locations for the tree brushes.
  The paths in the lower halves of each display represent the average value
  across different bacteria rather than timepoints, as in all previous figures.
  Though the samples originally do not include any hierarchical structure,
  hierarchical clustering provides such a structure which can then be
  interpreted using treeboxes.}\label{fig:gptimebox}
\end{figure}

In Figure \ref{fig:gptimebox}, we compare two subclusters from the
hierarchical clustering tree, after zooming to a few of the bacteria
that distinguish between the clusters. In contrast to the figures displayed to
this point, we only print time series associated with observed samples -- the
leaves of the hierarchical clustering tree. This reduces visual artifacts that
can be created by plotting many similar internal nodes, and which can overwhelm
patterns occuring in the leaves, which are those of central interest. Upon
revisiting the original data, it becomes clear that the samples highlighted on
the left come from freshwater samples, while those on the right come from soil
and skin, and looking up taxonomic groups associated with the distinguishing
bacteria confirms this. For example, many of the species with high
abundances in the left figure come from order Oceanospirillales.
